---
title: "Individual website portfolio: Project 1"
author: "Andrea Gerards"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---



```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
```



# Climate change and temperature anomalies 


If we wanted to study climate change, we can find data on the *Combined Land-Surface Air and Sea-Surface Water Temperature Anomalies* in the Northern Hemisphere at [NASA's Goddard Institute for Space Studies](https://data.giss.nasa.gov/gistemp). The [tabular data of temperature anomalies can be found here](https://data.giss.nasa.gov/gistemp/tabledata_v3/NH.Ts+dSST.txt)

To define temperature anomalies you need to have a reference, or base, period which NASA clearly states that it is the period between 1951-1980.


```{r weather_data, cache=TRUE}

weather <- 
  read_csv("https://data.giss.nasa.gov/gistemp/tabledata_v3/NH.Ts+dSST.csv", 
           skip = 1, 
           na = "***")

```



```{r tidyweather}

tidyweather <- weather %>%
  select(Year:Dec) %>% 
  pivot_longer(cols=c('Jan':'Dec'),names_to="Month",values_to="delta")


```

## Plotting Information

Let us plot the data using a time-series scatter plot, and add a trendline. To do that, we first need to create a new variable called `date` in order to ensure that the `delta` values are plot chronologically. 


> In the following chunk of code, I used the  argument, which does not run a chunk of code; I did so that you can knit the document before tidying the data and creating a new dataframe `tidyweather`. When you actually want to run this code and knit your document, you must delete ` `, **not just here but in all chunks were ` ` appears.**


```{r scatter_plot,  }

tidyweather <- tidyweather %>%
  mutate(date = ymd(paste(as.character(Year), Month, "1")),
         month = month(date, label=TRUE),
         year = year(date))

ggplot(tidyweather, aes(x=date, y = delta))+
  geom_point()+
  geom_smooth(color="red") +
  theme_bw() +
  labs (
    title = "Weather Anomalies"
  )

```

Is the effect of increasing temperature more pronounced in some months?

```{r facet_wrap, echo=FALSE}

ggplot(tidyweather, aes(x=date, y = delta))+
  geom_point()+
  
  geom_smooth(color="red") +
  theme_bw() +
  labs (
    title = "Weather Anomalies"
  )+
  facet_wrap(~month)

```


It is sometimes useful to group data into different time periods to study historical data. For example, we often refer to decades such as 1970s, 1980s, 1990s etc. to refer to a period of time. NASA calcuialtes a temperature anomaly, as difference form the base periof of 1951-1980. The code below creates a new data frame called `comparison` that groups data in five time periods: 1881-1920, 1921-1950, 1951-1980, 1981-2010 and 2011-present.


```{r intervals,  }

comparison <- tidyweather %>% 
  filter(Year>= 1881) %>%     #remove years prior to 1881
  #create new variable 'interval', and assign values based on criteria below:
  mutate(interval = case_when(
    Year %in% c(1881:1920) ~ "1881-1920",
    Year %in% c(1921:1950) ~ "1921-1950",
    Year %in% c(1951:1980) ~ "1951-1980",
    Year %in% c(1981:2010) ~ "1981-2010",
    TRUE ~ "2011-present"
  ))

```


```{r density_plot,  }

ggplot(comparison, aes(x=delta, fill=interval))+
  geom_density(alpha=0.2) +   #density plot with tranparency set to 20%
  theme_bw() +                #theme
  labs (
    title = "Density Plot for Monthly Temperature Anomalies",
    y     = "Density"         #changing y-axis label to sentence case
  )

```

```{r averaging,  }

#creating yearly averages
average_annual_anomaly <- tidyweather %>% 
  group_by(Year) %>%   #grouping data by Year
  
  # creating summaries for mean delta 
  # use `na.rm=TRUE` to eliminate NA (not available) values 
  summarise(annual_average_delta = mean(delta, na.rm=TRUE)) 

#plotting the data:
ggplot(average_annual_anomaly, aes(x=Year, y= annual_average_delta))+
  geom_point() +
  
  #Fit the best fit line, using LOESS method
  geom_smooth() +
  
  #change to theme_bw() to have white background + black frame around plot
  theme_bw() +
  labs (
    title = "Average Yearly Anomaly",
    y     = "Average Annual Delta"
  )                         


```