---
title: "My personality traits"
author: "Andrea Gerards"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---



```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, include=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(janitor)
library(httr)
library(readxl)
library(vroom)
library(fmsb)
```



### Overview 

Most people have an intuitive idea of personality that is not far away from how social scientists understand it. Our families are where we typically first learn that people have a degree of consistency in how they interact with others, and later as we make friends we interpret other people’s behavior in terms of character, personal style or habit. Later in the world of work we hear and see that leaders and co-workers succeed or have problems because of the way they work with others.

Below you can find a spider chart of my main personality traits plotted after having completed a 100-question survey. The results have been compared to the ones of the LBS cohort. 

If you believe we can share similar traits, let's get in touch!


```{r personality_traits, echo=FALSE, cache=TRUE}

# Basic Radar chart

# Construct the data set
data <- data.frame(Emotional_Reactiveness = c(5, 0, 2.5, 2.6),
                   Extraversion = c(5, 0, 3.8, 3.6),
                   Openness_to_Experience = c(5, 0, 3.6, 3.5),
                   Agreeableness = c(5, 0, 3.3, 3.6),
                   Conscientiousness = c(5, 0, 4.3, 3.9),
                   row.names = c("max", "min", "Me", "Cohort")
                   )

# Define fill colors
colors_fill <- c(scales::alpha("gray", 0.1),
                 scales::alpha("gold", 0.1))

# Define line colors
colors_line <- c(scales::alpha("darkblue", 0.9),
                scales::alpha("lightblue", 0.9))

# Create plot
radarchart(data, 
           axistype=1,
           seg = 5,  # Number of axis segments
           title = "",
           pcol = colors_line,
           pfcol = colors_fill,
           plwd = 4,
           cglcol="grey", cglty=1, axislabcol="grey", caxislabels=seq(0,5,1), cglwd=0.8)

# Add a legend
legend(x=0.9, 
       y=1.2, 
       legend = rownames(data[-c(1,2),]), 
       bty = "n", pch=20 , col = colors_line, cex = 1.05, pt.cex = 2)

```

