---
title: "NYT Mask Use Acceptance"
author: "Andrea Gerards"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: zenburn
    theme: flatly
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---



```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries,echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(janitor)
library(skimr)
library(vroom)
library(mosaic)
library(here)

#remotes::install_github("kjhealy/socviz")
library(socviz)
```


## Introduction 

In some American neighborhoods, it’s hard to spot even one person outside without a face covering. In others, the odds of seeing many maskless people are quite high.

Public health officials believe that face coverings can substantially slow transmission of the coronavirus, which is spreading rapidly in many states. But face coverings work best if they are adopted widely, and that is not the case everywhere. The accompanying map shows the odds of whether, if you encountered five people in a given area, all of them would be wearing masks.

In this analysis, the data comes from a large number of interviews conducted by the global data and survey firm Dynata at the request of The New York Times. The firm asked a question about mask use to obtain 250,000 survey responses between July 2 and July 14, enough data to provide estimates more detailed than the state level. (Several states have imposed new mask requirements since the completion of these interviews.)


```{r include=FALSE}
#Source for data url <- "https://github.com/nytimes/covid-19-data/raw/master/mask-use/mask-use-by-county.csv"

nyt_mask_survey <- read_csv(here::here("data", "nyt_mask_survey.csv"))

nyt_mask_survey <- nyt_mask_survey %>%
  clean_names() %>% 
  mutate(
    mostly_yes= frequently+always,
    mostly_no = never+rarely,
    delta = mostly_yes-mostly_no
  )

glimpse(nyt_mask_survey)
```

## Choropleth map

The FIPS code is a federal code that numbers states and territories of the US. It extends to the county level with an additional four digits, so every county in the US has a unique six-digit identifier, where the first two digits represent the state. 

I have used Kieran Healy's `socviz` package which among other things contains `county_map` and `county_data`  

```{r include=FALSE}
# America’s choropleths; use county_map that has all polygons 
# and county data with demographics/election data from socviz datafile
# The id field is the FIPS code for the county
county_map %>% 
  sample_n(5)

county_data %>%
    sample_n(5)


glimpse(county_data)
# we have data on 3195 FIPS....

glimpse(county_map)
# ... but to create a map, we translate these 3195 counties to 191,382 polygons!




```



```{r include=FALSE}


county_full <- left_join(county_map, county_data, by = "id")

county_masks_full <- left_join(county_full, nyt_mask_survey, 
                                by = c("id"="countyfp"))
```


## Building the choropleth plot

```{r echo=FALSE, plot_choropleths, cache=TRUE}
p <- ggplot(data = county_masks_full,
            mapping = aes(x = long, y = lat,
                          fill = delta, 
                          group = group))+
  geom_polygon(color = "gray90", size = 0.05) + 
  coord_equal()+
  scale_fill_gradientn(colours = c('#8c510a','#d8b365','#f6e8c3','#c7eae5','#5ab4ac','#01665e'))

p + labs(fill = "Mask acceptance, (Mostly Yes - Mostly No)", 
          caption = "“Estimates from The New York Times, based on roughly 250,000 interviews \nconducted by Dynata from July 2 to July 14, 2020”") +
  guides(fill = guide_legend(nrow = 1)) + 
  theme_map() + 
  theme(legend.position = "bottom")

```

The map shows broad regional patterns: Mask use is high in the Northeast and the West, and lower in the Plains and parts of the South. But it also shows many fine-grained local differences. Masks are widely worn in the District of Columbia, but there are sections of the suburbs in both Maryland and Virginia where norms seem to be different. In St. Louis and its western suburbs, mask use seems to be high. But across the Missouri River, it falls.



## Checking for relationships

Does mask use acceptance have any relation with some demographics? Let's explore the relationship between country household income, population, and % who voted republican in 2016

```{r echo=FALSE, ggpairs, cache=TRUE}
county_masks_full %>% 
  select(hh_income, pop, per_gop_2016, delta) %>% 
  GGally::ggpairs()+
  theme_minimal()
```
