---
title: "Voting in the UN General Assembly"
author: "Andrea Gerards"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---

```{r load-packages, include=FALSE}
list.of.packages <- c("unvotes", "tidyverse","lubridate", "DT") #check if packages are installed
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages, repos=c("http://rstudio.org/_packages", "http://cran.rstudio.com",dependencies=TRUE))

library(unvotes)
library(tidyverse)
library(lubridate)
library(DT)
library(ggthemes)
```

### Overview

Every September, leaders from around the world gather at the United Nations in New York to debate and pass resolutions on the biggest issues facing the planet.

Established after World War II, the UN has grown from 51 members in 1945 to 193 members in 2019. It also has two non-member observer states, Palestine and The Holy See.

Roll-call voting in the United Nations General Assembly (UNGA) has long attracted the attention of scholars; first to study the formation of voting blocs in the UNGA and more recently to create indicators for the common interests of states. 

The results of the analysis, that focuses primarily on western European contries as Italy, Germany and the UK as well as the US, highlights clear historical voting trends. In particular, visualizing the "Yes" votes to resolutions, the US show a constantly decreasing percentage of positive votes. This seems to be in line with its politcal strategical evolution. 
Digging into the data and grouping the votes in related issues, the US seems to be in line with European countries (Italy used as a proxy). Potential driver of the trend could be the presence of both within the NATO agreement.  

```{r view-un_votes, include=FALSE}
un_votes
```


```{r view-un_roll_calls, include=FALSE}
un_roll_calls
```


```{r view-un_roll_call_issues, include=FALSE}
un_roll_call_issues
```



```{r country-list, include=FALSE}
country_list <- c("United States of America", "China", 
                  "United Kingdom of Great Britain and Northern Ireland", "Italy", "Germany")
```


```{r plot-yearly-yes, echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}
un_votes %>%
  filter(country %in% country_list) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  group_by(year = year(date), country) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes")
    ) %>%
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_line() +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      subtitle = "1946 to 2015",
      y = "% Yes",
      x = "Year",
      color = "Country"
    )+
  theme_economist() + 
  scale_color_economist()
```

```{r plot-yearly-yes-issue, echo=FALSE, fig.width=10, message=FALSE, warning=FALSE}
un_votes %>%
  filter(country %in% c("United States of America", "Italy")) %>%
  inner_join(un_roll_calls, by = "rcid") %>%
  inner_join(un_roll_call_issues, by = "rcid") %>%
  group_by(country, year = year(date), issue) %>%
  summarize(
    votes = n(),
    percent_yes = mean(vote == "yes")
    ) %>%
  filter(votes > 5) %>%  # only use records where there are more than 5 votes
  ggplot(mapping = aes(x = year, y = percent_yes, color = country)) +
    geom_point() +
    geom_smooth(method = "loess", se = FALSE) +
    facet_wrap(~ issue) +
    labs(
      title = "Percentage of 'Yes' votes in the UN General Assembly",
      subtitle = "1946 to 2015",
      y = "% Yes",
      x = "Year",
      color = "Country"
    )+
  theme_economist() + 
  scale_color_economist()
  
```


## References

1. David Robinson (2017). unvotes: United Nations General Assembly Voting Data. 
R package version 0.2.0. https://CRAN.R-project.org/package=unvotes.
2. Erik Voeten "Data and Analyses of Voting in the UN General Assembly" 
Routledge Handbook of International Organization, edited by Bob Reinalda 
(published May 27, 2013).
3. Much of the analysis has been modeled on the examples presented in the 
[unvotes package vignette](https://cran.r-project.org/web/packages/unvotes/vignettes/unvotes.html).

```{r list-countries, include=FALSE}
un_votes %>% 
  select(country) %>%
  arrange(country) %>% 
  distinct() %>%
  datatable()
```
